% Lexical structure

%%\vfill\eject
\chapter{Lexical conventions}

This section gives an informal account of some of the lexical
conventions used in writing Scheme programs.  For a formal syntax of
Scheme, see section~\ref{BNF}.

\section{Identifiers}
\label{syntaxsection}

An identifier\mainindex{identifier} is any sequence of letters, digits, and
``extended identifier characters'' provided that it does not have a prefix
which is a valid number.  
However, the  \ide{.} token (a single period) used in the list syntax
is not an identifier.

All implementations of Scheme must support the following extended identifier
characters:

\begin{scheme}
!\ \$ \% \verb"&" * + - . / :\ < = > ? @ \verb"^" \verb"_" \verb"~" %
\end{scheme}

In addition, any character supported by an implementation can be used within an identifier when
specified using an \meta{inline hex escape}.  For example, the
identifier \verb|H\x65;llo| is the same as the identifier
\verb|Hello|, and in an implementation which supports the appropriate
Unicode character the identifier \verb|\x3BB;| is the same as the
identifier $\lambda$.

As a convenience, identifiers may also be written as a sequence of
zero or more
characters enclosed within vertical bars ({\cf $|$}), analogous to
string literals.  Any character, including whitespace characters, but
excluding the backslash and vertical bar characters,
may appear verbatim in such an identifier.
It is also possible to include the backslash and vertical bar characters,
as well as any other character,
in the identifier with an \meta{inline hex escape}.  Thus the
identifier \verb+|foo bar|+ is the same as the identifier
\verb+foo\x20;bar+.
Note that \verb+||+ is a valid identifier that is not equal to any other
identifier.

Here are some examples of identifiers:

\begin{scheme}
lambda                   q
list->vector             +soup+
{+}                        V17a
<=?                      a34kTMNs
->string                 ...
|two words|              two\backwhack{}x20;words
the-word-recursion-has-many-meanings%
\end{scheme}

See section~\ref{extendedalphas} for the formal syntax of identifiers.

\vest Identifiers have two uses within Scheme programs:
\begin{itemize}
\item Any identifier may be used as a variable\index{variable}
 or as a syntactic keyword\index{syntactic keyword}
(see sections~\ref{variablesection} and~\ref{macrosection}).

\item When an identifier appears as a literal or within a literal
(see section~\ref{quote}), it is being used to denote a {\em symbol}
(see section~\ref{symbolsection}).
\end{itemize}

In contrast with earlier revisions of the report~\cite{R5RS}, the
syntax distinguishes between upper and lower case in
identifiers and in characters specified via their names,
but not in numbers,
nor in \meta{inline hex escapes} used
in the syntax of identifiers, characters, or strings.
None of the identifiers defined in this report contain upper-case
characters, even when they may appear to do so as a result
of the English-language convention of capitalizing the word at
the beginning of a sentence.

The following directives are available to explicitly control case
folding.

\begin{entry}{%
{\cf{}\#!fold-case}\sharpbangindex{fold-case}\\
{\cf{}\#!no-fold-case}\sharpbangindex{no-fold-case}}

These directives may appear anywhere comments are permitted (see
section~\ref{wscommentsection})
and are treated as comments, except that they affect the reading
of subsequent data. The {\cf{}\#!fold-case} directive causes
the {\cf read} procedure to case-fold (as if by {\cf string-foldcase};
see section~\ref{stringsection}) each identifier and character name
subsequently read from the same port. The {\cf{}\#!no-fold-case} directive
causes the {\cf read} procedure to return to the default, non-folding behavior.
\end{entry}



\section{Whitespace and comments}
\label{wscommentsection}

\defining{Whitespace} characters include the space and newline characters.
(Implementations may provide additional whitespace characters such
as tab or page break.)  Whitespace is used for improved readability and
as necessary to separate tokens from each other, a token being an
indivisible lexical unit such as an identifier or number, but is
otherwise insignificant.  Whitespace can occur between any two tokens,
but not within a token.  Whitespace occurring inside a string
or inside a symbol delimited by vertical bars
is significant.

The lexical syntax includes several comment forms.  
Comments are treated exactly like whitespace.

A semicolon ({\tt;}) indicates the start of a line
comment.\mainindex{comment}\mainschindex{;}  The comment continues to the
end of the line on which the semicolon appears.  Comments are invisible
to Scheme, but the end of the line is visible as whitespace.  This
prevents a comment from appearing in the middle of an identifier or
number.

Another way to indicate a comment is to prefix a \hyper{datum}
(cf.\ section~\ref{datumsyntax}) with {\tt \#;}\sharpindex{;} as well as possible
\meta{whitespace} before the \hyper{datum}.  The comment consists of
the comment prefix {\tt \#;}, the space, and the \hyper{datum} together.  This
notation is useful for ``commenting out'' sections of code.

Block comments are indicated with properly nested {\tt
  \#|}\index{#"|@\texttt{\sharpsign\verticalbar}}\index{"|#@\texttt{\verticalbar\sharpsign}}
and {\tt |\#} pairs.

\begin{scheme}
\#|
   The FACT procedure computes the factorial
   of a non-negative integer.
|\#
(define fact
  (lambda (n)
    (if (= n 0)
        \#;(= n 1)
        1        ;Base case: return 1
        (* n (fact (- n 1))))))%
\end{scheme}


\section{Other notations}

\todo{Rewrite?}

For a description of the notations used for numbers, see
section~\ref{numbersection}.

\begin{description}{}{}

\item[{\tt.\ + -}]
These are used in numbers, and may also occur anywhere in an identifier.
A delimited plus or minus sign by itself
is also an identifier.
A delimited period (not occurring within a number or identifier) is used
in the notation for pairs (section~\ref{listsection}), and to indicate a
rest-parameter in a  formal parameter list (section~\ref{lambda}).
Note that a sequence of two or more periods {\em is} an identifier.

\item[\tt( )]
Parentheses are used for grouping and to notate lists
(section~\ref{listsection}).

\item[\singlequote]
The single quote character is used to indicate literal data (section~\ref{quote}).

\item[\backquote]
The backquote character is used to indicate partly-constant
data (section~\ref{quasiquote}).

\item[\tt, ,@]
The character comma and the sequence comma at-sign are used in conjunction
with backquote (section~\ref{quasiquote}).

\item[\tt"]
The double quote character is used to delimit strings (section~\ref{stringsection}).

\item[\backwhack]
Backslash is used in the syntax for character constants
(section~\ref{charactersection}) and as an escape character within string
constants (section~\ref{stringsection}) and identifiers
(section~\ref{extendedalphas}).

% A box used because \verb is not allowed in command arguments.
\setbox0\hbox{\tt \verb"[" \verb"]" \verb"{" \verb"}"}
\item[\copy0]
Left and right square brackets and curly braces
are reserved for possible future extensions to the language.

\item[\sharpsign] Sharp sign is used for a variety of purposes depending on
the character that immediately follows it:

\item[\schtrue{} \schfalse{}]
These are the boolean constants (section~\ref{booleansection}),
along with the alternatives \sharpfoo{true} and \sharpfoo{false}.

\item[\sharpsign\backwhack]
This introduces a character constant (section~\ref{charactersection}).

\item[\sharpsign\tt(]
This introduces a vector constant (section~\ref{vectorsection}).  Vector constants
are terminated by~{\tt)}~.

\item[\sharpsign\tt u8(]
This introduces a bytevector constant (section~\ref{bytevectorsection}).  Bytevector constants
are terminated by~{\tt)}~.

\item[{\tt\#e \#i \#b \#o \#d \#x}]
These are used in the notation for numbers (section~\ref{numbernotations}).

\item[\tt{\#\hyper{n}= \#\hyper{n}\#}]
These are used for labeling and referencing other literal data (section~\ref{labelsection}).

\end{description}

\section{Datum labels}\unsection
\label{labelsection}

\begin{entry}{%
\pproto{\#\hyper{n}=\hyper{datum}}{lexical syntax}
\pproto{\#\hyper{n}\#}{lexical syntax}}

The lexical syntax
\texttt{\#\hyper{n}=\hyper{datum}} reads the same as \hyper{datum}, but also
results in \hyper{datum} being labelled by \hyper{n}.
It is an error if \hyper{n} is not a sequence of digits.

The lexical syntax \texttt{\#\hyper{n}\#} serves as a reference to some
object labelled by \texttt{\#\hyper{n}=}; the result is the same
object as the \texttt{\#\hyper{n}}= as compared with {\cf eqv?}
(see section~\ref{equivalencesection}). 

Together, these syntaxes permit the notation of
structures with shared or circular substructure.

\begin{scheme}
(let ((x (list 'a 'b 'c)))
  (set-cdr! (cddr x) x)
  x)                       \ev \#0=(a b c . \#0\#)
\end{scheme}

The scope of a datum label is the portion of the datum in which it appears
that is to the right of the label.
Consequently, a reference \texttt{\#\hyper{n}\#} may occur only after a label
\texttt{\#\hyper{n}=}; it is an error to attempt a forward reference.  In
addition, it is an error if the reference appears as the labelled object itself
(as in \texttt{\#\hyper{n}= \#\hyper{n}\#}),
because the object labelled by \texttt{\#\hyper{n}=} is not well
defined in this case.

It is an error for a \hyper{program} or \hyper{library} to include
circular references.  In particular,
it is an error for {\cf quasiquote} (section~\ref{quasiquote}) to contain them.

\begin{scheme}
\#1=(begin (display \#\backwhack{}x) . \#1\#)
                       \ev \scherror%
\end{scheme}
\end{entry}

